---
title: "Process Data Analysis"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# MAC
knitr::opts_knit$set(root.dir = '../Data')

```


# Import Libraries
```{r, results = "hide", message=FALSE}

library(tidyverse)
library(dplyr)
library(tidyr)
library(caret)
library(glmnet)
library(grid)
library(pROC)
library(umap)
library(dbscan)
library(pdp)
library(BBmisc)
library(ggplot2)
library(cowplot)

```


# Import Data
```{r, results = "hide", message=FALSE}

sc <- read.csv('sc.csv')
df_all <- read.csv('df_all.csv')

```

# Define cross validation function

```{r}

nested_cv_reg <- function(data, target, inner_method = 'cv', inner_number = 10,
                      outer_method = 'kfold', outer_number = 10,
                      p_mc = 0.75,
                      random_seed=NULL){
  
  
  # cross validation strategy for inner loop
  ctrl <- trainControl(method = "cv", number = inner_number)
  
  # parameter grid for random forest
  rf_grid <- data.frame(.mtry = seq(2, (ncol(data) * 0.8), 2))
  
  
  # create empty list for results
  cv_results <- list('imp' = NULL, 'rf' = NULL)
  
  
  # create folds for outer k-fold cv
  set.seed(random_seed)
  kfolds <- rep(1:outer_number, length.out = nrow(data)) %>% sample(replace = F)

  # nested cross validation loop 
  for (i in 1:outer_number){
    
    print(paste('Outer loop:', i))
    
    # set random seed 
    set.seed(i)
    
    # train test split outer loop
    if (outer_method == 'monte_carlo'){
    in_train <- createDataPartition(y = data[[target]],
                                  p = p_mc, list = FALSE)
    }
    
    if (outer_method == 'kfold'){
    in_train <- which(kfolds != i)
    }
  
    training <- data[ c(in_train), ]
    testing  <- data[-c(in_train), ]
    
    X_train <- training %>% dplyr::select(-all_of(target))
    y_train <- training[[target]] 
    
    X_test <- testing %>% dplyr::select(-all_of(target))
    y_test <- testing[[target]]
  
  
    #### random forest
    # parameter tuning in inner loop
    mdl_rf_inner <- caret::train(X_train, y_train, data = training, method = "rf",
                                 preProcess = c("center", "scale"),
                                 ntrees = 1000, importance = TRUE,
                                 tuneGrid = rf_grid,
                                 trControl = ctrl)
    
    
    # get feature importance
    feat_imp <- caret::varImp(mdl_rf_inner, scale=F)
    feat_imp <- feat_imp$importance %>% t()
    cv_results[[1]] <-  cv_results[[1]] %>% rbind(feat_imp)

    
    # evaluate on test set in outer loop 
    y_pred_rf <- predict(mdl_rf_inner, newdata = X_test) %>% scale()
    y_test <- y_test %>% scale()
    
    eval_rf <- defaultSummary(data.frame(obs = y_test, pred = y_pred_rf))
    cor_acc <- cor(y_test, y_pred_rf)
    
    rss <- sum((y_pred_rf - y_test) ^ 2)
    tss <- sum((y_test - mean(y_test)) ^ 2)
    rsq <- 1 - rss/tss
    
    eval_rf = eval_rf %>% 
      as.data.frame() %>% t() %>% 
      cbind(rsq) %>% 
      cbind(cor_acc)
    
    rss <- NULL
    tss <- NULL
    rsq <- NULL

    cv_results[[2]] <-  cv_results[[2]] %>% rbind(eval_rf)
    
  }
  
  return(cv_results)

}

```


# Exctract Process Features - SC

```{r}
variables = c('id', 'item', 'time_taken', 'distance_travelled', 
         'xPos', 'yPos', 'zPos', 'Pitch', 'Yaw', 'ray.x.', 'ray.y.', 
         'ray.z.', 'ray.distance.', 'distance_goal', 'looking_model', 
         'looking_goal', 'steps', 'correctBlocks', 'incorrectBlocks', 
         'answer')

features = c('time_taken', 'distance_travelled', 
         'xPos', 'yPos', 'zPos', 'Pitch', 'Yaw', 'ray.x.', 'ray.y.', 
         'ray.z.', 'ray.distance.', 'distance_goal', 'looking_model', 
         'looking_goal', 'steps', 'correctBlocks', 'incorrectBlocks')
```

```{r, warning=FALSE, message=FALSE}
sc_feat <- sc %>% 
  dplyr::select(all_of(variables)) %>% 
  mutate_at(features, as.character) %>% 
  mutate_at(features, as.numeric)


sc_feat <- sc_feat %>% 
  filter(item != 1 & item != 2) %>%
  dplyr::group_by(id) %>%
  dplyr::summarize(distance_travelled = max(as.numeric(distance_travelled), na.rm = T),
            xPos_range = diff(range(as.numeric(xPos), na.rm = T)), 
            xPos_mean = mean(as.numeric(xPos), na.rm = T), 
            xPos_sd = sd(as.numeric(xPos), na.rm = T), 
            yPos_range = diff(range(as.numeric(yPos), na.rm = T)), 
            yPos_mean = mean(as.numeric(yPos), na.rm = T), 
            yPos_sd = sd(as.numeric(yPos), na.rm = T), 
            zPos_range = diff(range(as.numeric(zPos), na.rm = T)),
            zPos_mean = mean(as.numeric(zPos), na.rm = T), 
            zPos_sd = sd(as.numeric(zPos), na.rm = T), 
            pitch_range = diff(range(as.numeric(Pitch), na.rm = T)),
            pitch_mean = mean(as.numeric(Pitch), na.rm = T), 
            pitch_sd = sd(as.numeric(Pitch), na.rm = T), 
            yaw_range = diff(range(as.numeric(Yaw), na.rm = T)),
            yaw_mean = mean(as.numeric(Yaw), na.rm = T),
            yaw_sd = sd(as.numeric(Yaw), na.rm = T),
            ray_x_range = diff(range(as.numeric(ray.x.), na.rm = T)),
            ray_x_mean = mean(as.numeric(ray.x.), na.rm = T),
            ray_x_sd = sd(as.numeric(ray.x.), na.rm = T),
            ray_y_range = diff(range(as.numeric(ray.y.), na.rm = T)),
            ray_z_mean = mean(as.numeric(ray.z.), na.rm = T),
            ray_z_sd = sd(as.numeric(ray.z.), na.rm = T),
            ray_dist_range = diff(range(as.numeric(ray.distance.), na.rm = T)),
            ray_dist_mean = mean(as.numeric(ray.distance.), na.rm = T), 
            ray_dist_sd = sd(as.numeric(ray.distance.), na.rm = T), 
            looking_model = max(as.numeric(looking_model), na.rm = T),
            looking_goal = max(as.numeric(looking_goal), na.rm = T),
            looking_ratio = looking_model/looking_goal) %>%
  ungroup()

dim(sc_feat)
sc_feat
```


# Predict Scores
### Merge data
```{r}
sc_feat_all <- sc_feat %>% inner_join(df_all, by="id") 
sc_feat_all
```


### Predict SC Scores From Process Data
```{r}

sc_feat_all_pred_sc <- sc_feat_all %>% 
  dplyr::select(-c(pc, mr, VKMR, SPM, xPos_range, ray_x_range)) %>% 
  column_to_rownames("id")

cv_results_sc <- nested_cv_reg(data = sc_feat_all_pred_sc, 
                               target = 'sc',
                               outer_method = 'monte_carlo',
                               inner_number = 10, 
                               outer_number = 30,
                               p_mc = 0.85,
                               random_seed = 1)

```

### Predict MR Scores From Process Data
```{r}

sc_feat_all_pred_mr <- sc_feat_all %>% 
  dplyr::select(-c(pc, sc, SPM, VKMR, xPos_range, ray_x_range)) %>% 
  column_to_rownames("id")

cv_results_mr <- nested_cv_reg(data = sc_feat_all_pred_mr, 
                           target = 'mr',
                           #outer_method = 'monte_carlo',
                           inner_number = 10, 
                           outer_number = 30,
                           p_mc = 0.85,
                           random_seed = 1)

```

### Predict PC Scores From Process Data
```{r}

sc_feat_all_pred_pc <- sc_feat_all %>% 
  dplyr::select(-c(mr, sc, SPM, VKMR, xPos_range, ray_x_range)) %>% 
  column_to_rownames("id")

cv_results_pc <- nested_cv_reg(data = sc_feat_all_pred_pc, 
                               target = 'pc',
                               outer_method = 'monte_carlo',
                               inner_number = 10, 
                               outer_number = 30,
                               p_mc = 0.85,
                               random_seed = 1)

```


### Predict VKMR Scores From Process Data
```{r}

sc_feat_all_pred_vkmr <- sc_feat_all %>% 
  dplyr::select(-c(pc, mr, sc, SPM, xPos_range, ray_x_range)) %>% 
  column_to_rownames("id")

cv_results_vkmr <- nested_cv_reg(data = sc_feat_all_pred_vkmr,
                                 target = 'VKMR',
                                 outer_method = 'monte_carlo',
                                 inner_number = 10, 
                                 outer_number = 30,
                                 p_mc = 0.85,
                                 random_seed = 1)

```


### Predict SPM Scores From Process Data
```{r}

sc_feat_all_pred_spm <- sc_feat_all %>%
  dplyr::select(-c(pc, mr, sc, VKMR, xPos_range, ray_x_range)) %>%
  column_to_rownames("id")

cv_results_spm <- nested_cv_reg(data = sc_feat_all_pred_spm,
                                target = 'SPM',
                                outer_method = 'monte_carlo',
                                inner_number = 10,
                                outer_number = 30,
                                p_mc = 0.85,
                                random_seed = 1)

```


### Aggregate results
```{r}

results <- list(cv_results_sc, cv_results_pc, cv_results_mr, cv_results_vkmr, cv_results_spm)
targets <- c('SC', 'PC', 'MR', 'VKMR', 'SPM')

rf_metrics <- NULL

for(i in results){

  rf_mean <- i$rf %>% as.data.frame() %>% 
    replace( is.na(.), 0) %>% map_dbl(mean)
  names(rf_mean) <- paste0(names(rf_mean), '_mean')
  rf_se <- i$rf %>% as.data.frame() %>% 
    replace( is.na(.), 0) %>% map_dbl(function(x) sd(x)/sqrt(length(x)))
  names(rf_se) <- paste0(names(rf_se), '_se')
  
  rf_metrics <- rf_metrics %>% rbind(c(rf_mean, rf_se))
}

rownames(rf_metrics) <-targets
rf_metrics %>% as.data.frame %>% 
  select(RMSE_mean, RMSE_se, MAE_mean, MAE_se, V5_mean, V5_se)


rf_metrics %>% as.data.frame %>% 
  select(RMSE_mean, RMSE_se, MAE_mean, MAE_se, V5_mean, V5_se) %>% 
  xtable::xtable()

```


# Analyze feature importances

```{r}
feat_imps_sc <- cv_results_sc$imp %>% as.data.frame() %>% map_dbl(mean)
feat_imps_sc_sd <- cv_results_sc$imp %>% as.data.frame() %>% map_dbl(sd)

feat_imps_sc <- feat_imps_sc[names(feat_imps_sc) != 'ray_y_range']
feat_imps_sc_sd <- feat_imps_sc_sd[names(feat_imps_sc_sd) != 'ray_y_range']

feat_imps_sc_sd <- feat_imps_sc_sd[order(feat_imps_sc, decreasing = T)]
feat_imps_sc <- feat_imps_sc[order(feat_imps_sc, decreasing = T)]

feat_imps_sc %>% barplot(ylim=c(-1,15), las = 2)
```


```{r}
predictors <- names(sc_feat_all[2:29])
predictors

```


```{r}

sc_feat_all <- sc_feat_all %>% select(-ray_y_range)

sc_feat_norm <- sc_feat_all[-1] %>% 
  normalize(method='range',range=c(0,1)) %>% 
  as.data.frame()
```

```{r}
cbind(feat_imps_sc, feat_imps_sc_sd) %>% as.data.frame() %>% xtable::xtable()
```


```{r}
ggplot(sc_feat_norm, aes(y = sc)) + 
  geom_smooth(aes(x = yPos_mean, col='yPos_mean'), se=F) +
  geom_smooth(aes(x = yPos_sd, col='yPos_sd'), se=F) +
  geom_smooth(aes(x = distance_travelled, col='distance_travelled'), se=F) +
  geom_smooth(aes(x = pitch_range, col='pitch_range'), se=F) +
  geom_smooth(aes(x = ray_dist_mean, col='ray_dist_mean'), se=F) +
  ylab('Normalized SC Scores') +
  xlab('Normalized Feature Values') +
  #scale_color_manual(labels = c("T999", "T888","T99", "T88", "T77"), values = c('1', '2', '3', '4', '5')) +
  labs(color='Features') +
  theme_bw()

  

```

```{r}
ggplot(sc_feat_norm, aes(y = VKMR)) + 
  geom_smooth(aes(x = yPos_mean, col='yPos_mean'), se=F) +
  geom_smooth(aes(x = yPos_sd, col='yPos_sd'), se=F) +
  geom_smooth(aes(x = distance_travelled, col='distance_travelled'), se=F) +
  geom_smooth(aes(x = pitch_range, col='pitch_range'), se=F) +
  geom_smooth(aes(x = ray_dist_mean, col='ray_dist_mean'), se=F) +
  ylab('Normalized SC Scores') +
  xlab('Normalized Feature Values') +
  #scale_color_manual(labels = c("T999", "T888","T99", "T88", "T77"), values = c('1', '2', '3', '4', '5')) +
  labs(color='Features') +
  theme_bw()


```





```{r}

plots <- list()
for (i in names(feat_imps_sc)){
  gg <- ggplot(sc_feat_norm, aes_string(x = i, y = sc_feat_norm$sc)) +
    geom_smooth() +
  ylab('')
  #print(gg)
  plots[[i]] <- as_grob(gg)
}


grid.arrange(plots[[1]], plots[[2]],vplots[[3]], plots[[4]], plots[[5]], 
             plots[[6]], plots[[7]], plots[[8]], plots[[9]], plots[[10]], 
             plots[[11]], plots[[12]], plots[[13]], plots[[14]], plots[[15]],
             plots[[16]], plots[[17]], plots[[18]], plots[[19]], plots[[20]],
             plots[[21]], plots[[22]], plots[[23]], plots[[24]], plots[[25]], 
             ncol=5
             )
```



# Cluster Analysis
### Identify clusters using UMAP and DBSCAN
```{r}
# select only features
sc_feat_all_feats <- sc_feat_all %>% 
  dplyr::select(distance_travelled:looking_ratio)

# project features to umap space
umap_config <- umap.defaults
umap_config$random_state <- 123

sc_feat_umap <- umap(sc_feat_all_feats, umap_config, verbose=T)
sc_feat_umap

# run dbscan cluster analysis
dbscan_cluster = dbscan(sc_feat_umap$layout, eps=3)
cluster <- dbscan_cluster$cluster

sc_feat_all_clust <- sc_feat_all %>% 
  cbind(cluster)

df_plot <- data.frame(sc_feat_umap$layout, as.factor(cluster))
names(df_plot) <- c('Dimension_1', 'Dimension_2', 'cluster')

ggplot(data = df_plot, 
       aes(x = Dimension_1, y = Dimension_2, color = cluster, size = 1.5, alpha = 0.5)) +
  geom_point() +
  guides(alpha = FALSE) +
  guides(size = FALSE) +
  theme_bw()
  
```

### Test for cluster differences 
```{r}
sc_feat_all_clust %>% 
  dplyr::group_by(cluster) %>% 
  dplyr::summarise(mean(sc))
```


```{r}
anova_results <- aov(sc~as.factor(cluster), data = sc_feat_all_clust)
anova_results %>% summary()
lm(sc~as.factor(cluster), data = sc_feat_all_clust) %>% summary()

# Pairwise t-test
pwtt <- pairwise.t.test(sc_feat_all_clust$sc, 
                sc_feat_all_clust$cluster, 
                p.adjust.method ='none')

pwtt$p.value %>% xtable()

```


# Analyze cluster differences
```{r}
# normalize predictors
sc_feat_all_clust[predictors] <- sc_feat_all_clust[predictors] %>% 
  normalize(method='range',range=c(0,1))

# cluster means
cluster_means <- sc_feat_all_clust %>% 
  group_by(cluster) %>% 
  summarise_at(predictors, mean)
```


```{r}
# compare clusters across predictors
for (i in 1:5) {
  cluster_means %>% t() %>% .[-1,] %>% .[,i] %>% barplot()
}
```


```{r}
# compare predictors across clusters
for (i in predictors) {
  x <- cluster_means %>% .[,-1] %>% .[i] %>% t() 
  names(x) <- 1:5
  x %>% barplot(main = i)
}

```


