---
title: "Process Data Analysis"
author: "Heinrich Peters"
date: "4/10/2020"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# MAC
knitr::opts_knit$set(root.dir = '../Data')

```


# Import Libraries
```{r, results = "hide", message=FALSE}

library(tidyverse)
library(dplyr)
library(tidyr)
library(caret)
library(glmnet)
library(grid)
library(pROC)
library(umap)
library(dbscan)

```


# Import Data
```{r, results = "hide", message=FALSE}

# pc <- read.csv('pc.csv')
# mr <- read.csv('mr.csv')
sc <- read.csv('sc.csv')
df_all <- read.csv('df_all.csv')

```

# Define cross validation functions
```{r}

nested_cv_clf <- function(data, target, inner_method = 'cv', inner_number = 10,
                      outer_method = 'kfold', outer_number = 10, p_mc = 0.75,
                      imb_sampling = NULL,
                      random_seed = NULL){
            
  # cross validation strategy for inner loop
  ctrl <- trainControl(method = "cv", number = inner_number,
                       sampling = imb_sampling)
  
  # parameter grid for elastic net
  enet_grid <- expand.grid(.lambda = seq(.05, 1, length = 5),
                      .alpha = seq(.05, 1, length = 5))
  
  # parameter grid for random forest
  rf_grid <- data.frame(.mtry = seq(2, (ncol(data) * 0.8), 5))
  
  
  # create empty list for results
  cv_results <- list('enet' = NULL, 'rf' = NULL)
  
  # create folds for outer k-fold cv
  set.seed(random_seed)
  kfolds <- rep(1:outer_number, length.out = nrow(data)) %>% sample(replace = F)

  # nested cross validation loop 
  for (i in 1:outer_number){
    
    print(paste('Outer loop:', i))
    
    # set random seed 
    set.seed(i)
    
    # train test split outer loop
    if (outer_method == 'monte_carlo'){
    in_train <- createDataPartition(y = data[[target]],
                                  p = p_mc, list = FALSE)
    }
    
    if (outer_method == 'kfold'){
    in_train <- which(kfolds != i)
    }
  
    training <- data[ c(in_train), ]
    testing  <- data[-c(in_train), ]
    
    X_train <- training %>% dplyr::select(-all_of(target))
    y_train <- training[[target]]
    
    X_test <- testing %>% dplyr::select(-all_of(target))
    y_test <- testing[[target]]
    
    
    #### LASSO
    # parameter tuning in inner loop
    mdl_enet_inner <- caret::train(data.matrix(X_train), y_train, method = "glmnet",
                                   preProcess = c('nzv', "center", "scale"),
                                   trControl = ctrl, 
                                   tuneGrid = enet_grid)
    
    # evaluate on test set in outer loop 
    y_prob_enet <- predict(mdl_enet_inner, newdata = X_test, type = "prob")[,2]
    y_pred_enet <- predict(mdl_enet_inner, newdata = X_test)
    
    tryCatch({
        conf_mat_enet <- confusionMatrix(y_pred_enet, y_test)
        roc_enet <- roc(y_test, y_prob_enet)
        AUC <- auc(roc_enet)},
      error = function(cond){
        message('AUC was not computed')
        AUC = NA},
      finally = {})
    
    eval_enet <- cbind(t(data.frame(conf_mat_enet$overall)), 
                         t(data.frame(conf_mat_enet$byClass)), 
                         data.frame(AUC))
    
    cv_results[[1]] <-  cv_results[[1]] %>% rbind(eval_enet)
    
    #### random forest
    # parameter tuning in inner loop
    mdl_rf_inner <- caret::train(data.matrix(X_train), y_train, data = training, method = "rf",
                                 preProcess = c('nzv', "center", "scale"),
                                 ntrees = 1000, importance = TRUE,
                                 tuneGrid = rf_grid,
                                 trControl = ctrl)
    
    
    # evaluate on test set in outer loop 
    y_prob_rf <- predict(mdl_rf_inner, newdata = X_test, type = "prob")[,2]
    y_pred_rf <- predict(mdl_rf_inner, newdata = X_test)
    
    tryCatch({
        conf_mat_rf <- confusionMatrix(y_pred_rf, y_test)
        roc_rf <- roc(y_test, y_prob_rf)
        AUC <- auc(roc_rf)},
      error = function(cond){
        message('AUC was not computed')
        AUC = NA},
      finally = {})
    
    eval_rf <- cbind(t(data.frame(conf_mat_rf$overall)), 
                         t(data.frame(conf_mat_rf$byClass)), 
                         data.frame(AUC))
    
    cv_results[[2]] <-  cv_results[[2]] %>% rbind(eval_rf)
  }
  
  rownames(cv_results$enet) <- 1:outer_number
  rownames(cv_results$rf) <- 1:outer_number
  
  return(cv_results)

}

```


```{r}

nested_cv_reg <- function(data, target, inner_method = 'cv', inner_number = 10,
                      outer_method = 'kfold', outer_number = 10,
                      p_mc = 0.75,
                      random_seed=NULL){
  
  
  # cross validation strategy for inner loop
  ctrl <- trainControl(method = "cv", number = inner_number)
  
  # parameter grid for random forest
  rf_grid <- data.frame(.mtry = seq(2, (ncol(data) * 0.8), 5))
  
  
  # create empty list for results
  cv_results <- list('enet' = NULL, 'rf' = NULL)
  
  
  # create folds for outer k-fold cv
  set.seed(random_seed)
  kfolds <- rep(1:outer_number, length.out = nrow(data)) %>% sample(replace = F)

  # nested cross validation loop 
  for (i in 1:outer_number){
    
    print(paste('Outer loop:', i))
    
    # set random seed 
    set.seed(i)
    
    # train test split outer loop
    if (outer_method == 'monte_carlo'){
    in_train <- createDataPartition(y = data[[target]],
                                  p = p_mc, list = FALSE)
    }
    
    if (outer_method == 'kfold'){
    in_train <- which(kfolds != i)
    }
  
    training <- data[ c(in_train), ]
    testing  <- data[-c(in_train), ]
    
    X_train <- training %>% dplyr::select(-all_of(target))
    y_train <- training[[target]] 
    
    X_test <- testing %>% dplyr::select(-all_of(target))
    y_test <- testing[[target]]
  
  
    #### random forest
    # parameter tuning in inner loop
    mdl_rf_inner <- caret::train(X_train, y_train, data = training, method = "rf",
                                 preProcess = c("center", "scale"),
                                 ntrees = 1000, importance = TRUE,
                                 tuneGrid = rf_grid,
                                 trControl = ctrl)
    
    # evaluate on test set in outer loop 
    y_pred_rf <- predict(mdl_rf_inner, newdata = X_test)
    eval_rf <- defaultSummary(data.frame(obs = y_test, pred = y_pred_rf))
    cor_acc <- cor(y_test, y_pred_rf)
    
    rss <- sum((y_pred_rf - y_test) ^ 2)
    tss <- sum((y_test - mean(y_test)) ^ 2)
    rsq <- 1 - rss/tss
    
    eval_rf = eval_rf %>% 
      as.data.frame() %>% t() %>% 
      cbind(rsq) %>% 
      cbind(cor_acc)
    
    rss <- NULL
    tss <- NULL
    rsq <- NULL

    cv_results[[2]] <-  cv_results[[2]] %>% rbind(eval_rf)
    
  }
  
  return(cv_results)

}

```


# Exctract Process Features - SC

### Define list of variables for feature extraction
```{r}
variables = c('id', 'item', 'time_taken', 'distance_travelled', 
         'xPos', 'yPos', 'zPos', 'Pitch', 'Yaw', 'ray.x.', 'ray.y.', 
         'ray.z.', 'ray.distance.', 'distance_goal', 'looking_model', 
         'looking_goal', 'steps', 'correctBlocks', 'incorrectBlocks', 
         'answer')

features = c('time_taken', 'distance_travelled', 
         'xPos', 'yPos', 'zPos', 'Pitch', 'Yaw', 'ray.x.', 'ray.y.', 
         'ray.z.', 'ray.distance.', 'distance_goal', 'looking_model', 
         'looking_goal', 'steps', 'correctBlocks', 'incorrectBlocks')
```

### Extract features at person level
```{r}
sc_feat <- sc %>% 
  dplyr::select(all_of(variables)) %>% 
  mutate_at(features, as.character) %>% 
  mutate_at(features, as.numeric)


sc_feat <- sc_feat %>% 
  filter(item != 1 & item != 2) %>%
  dplyr::group_by(id) %>%
  dplyr::summarize(distance_travelled = max(as.numeric(distance_travelled), na.rm = T),
            xPos_range = diff(range(as.numeric(xPos), na.rm = T)), 
            xPos_mean = mean(as.numeric(xPos), na.rm = T), 
            xPos_sd = sd(as.numeric(xPos), na.rm = T), 
            yPos_range = diff(range(as.numeric(yPos), na.rm = T)), 
            yPos_mean = mean(as.numeric(yPos), na.rm = T), 
            yPos_sd = sd(as.numeric(yPos), na.rm = T), 
            zPos_range = diff(range(as.numeric(zPos), na.rm = T)),
            zPos_mean = mean(as.numeric(zPos), na.rm = T), 
            zPos_sd = sd(as.numeric(zPos), na.rm = T), 
            pitch_range = diff(range(as.numeric(Pitch), na.rm = T)),
            pitch_mean = mean(as.numeric(Pitch), na.rm = T), 
            pitch_sd = sd(as.numeric(Pitch), na.rm = T), 
            yaw_range = diff(range(as.numeric(Yaw), na.rm = T)),
            yaw_mean = mean(as.numeric(Yaw), na.rm = T),
            yaw_sd = sd(as.numeric(Yaw), na.rm = T),
            ray_x_range = diff(range(as.numeric(ray.x.), na.rm = T)),
            ray_x_mean = mean(as.numeric(ray.x.), na.rm = T),
            ray_x_sd = sd(as.numeric(ray.x.), na.rm = T),
            ray_y_range = diff(range(as.numeric(ray.y.), na.rm = T)),
            ray_z_mean = mean(as.numeric(ray.z.), na.rm = T),
            ray_z_sd = sd(as.numeric(ray.z.), na.rm = T),
            ray_dist_range = diff(range(as.numeric(ray.distance.), na.rm = T)),
            ray_dist_mean = mean(as.numeric(ray.distance.), na.rm = T), 
            ray_dist_sd = sd(as.numeric(ray.distance.), na.rm = T), 
            looking_model = max(as.numeric(looking_model), na.rm = T),
            looking_goal = max(as.numeric(looking_goal), na.rm = T),
            looking_ratio = looking_model/looking_goal) %>%
  ungroup()

dim(sc_feat)
sc_feat
```

### Extract features at item level
```{r}
sc_feat_item <- sc %>% 
  dplyr::select(all_of(variables)) %>% 
  mutate_at(features, as.character) %>% 
  mutate_at(features, as.numeric)

  
 sc_feat_item <- sc_feat_item %>% 
  filter(item != 1 & item != 2) %>%
  dplyr::group_by(id, item) %>%
  dplyr::summarise(distance_travelled = max(as.numeric(distance_travelled), na.rm = T),
            xPos_range = diff(range(as.numeric(xPos), na.rm = T)), 
            xPos_mean = mean(as.numeric(xPos), na.rm = T), 
            xPos_sd = sd(as.numeric(xPos), na.rm = T), 
            yPos_range = diff(range(as.numeric(yPos), na.rm = T)), 
            yPos_mean = mean(as.numeric(yPos), na.rm = T), 
            yPos_sd = sd(as.numeric(yPos), na.rm = T), 
            zPos_range = diff(range(as.numeric(zPos), na.rm = T)),
            zPos_mean = mean(as.numeric(zPos), na.rm = T), 
            zPos_sd = sd(as.numeric(zPos), na.rm = T), 
            pitch_range = diff(range(as.numeric(Pitch), na.rm = T)),
            pitch_mean = mean(as.numeric(Pitch), na.rm = T), 
            pitch_sd = sd(as.numeric(Pitch), na.rm = T), 
            yaw_range = diff(range(as.numeric(Yaw), na.rm = T)),
            yaw_mean = mean(as.numeric(Yaw), na.rm = T),
            yaw_sd = sd(as.numeric(Yaw), na.rm = T),
            ray_x_range = diff(range(as.numeric(ray.x.), na.rm = T)),
            ray_x_mean = mean(as.numeric(ray.x.), na.rm = T),
            ray_x_sd = sd(as.numeric(ray.x.), na.rm = T),
            ray_y_range = diff(range(as.numeric(ray.y.), na.rm = T)),
            ray_z_mean = mean(as.numeric(ray.z.), na.rm = T),
            ray_z_sd = sd(as.numeric(ray.z.), na.rm = T),
            ray_dist_range = diff(range(as.numeric(ray.distance.), na.rm = T)),
            ray_dist_mean = mean(as.numeric(ray.distance.), na.rm = T), 
            ray_dist_sd = sd(as.numeric(ray.distance.), na.rm = T), 
            looking_model = max(as.numeric(looking_model), na.rm = T),
            looking_goal = max(as.numeric(looking_goal), na.rm = T),
            looking_ratio = looking_model/looking_goal, 
            answer = last(answer)) %>% 
  ungroup()

dim(sc_feat_item)
sc_feat_item
```

### Spread variables by item
```{r}
# spread variables by item 
sc_feat_item_wide = sc_feat_item %>% dplyr::select(id) %>% unique()

for (i in names(sc_feat_item)[3:length(sc_feat_item)-1]){
  
  long_temp <- sc_feat_item %>% dplyr::select(id, item, all_of(i))
  wide_temp <- long_temp %>% tidyr::spread(key=item, value = i)
  names(wide_temp) <- paste("sc",names(wide_temp), i, sep='_')
  names(wide_temp)[1] <- "id"
  sc_feat_item_wide = sc_feat_item_wide %>% inner_join(wide_temp, by="id")
}

sc_feat_item_wide <- sc_feat_item_wide[!str_detect(names(sc_feat_item_wide), 'item')]

sc_feat_item_wide
```

# Classificatio task - predicting inividual item solutions
## Formatting
```{R}

sc_feat_item_clean <- sc_feat_item %>% mutate(id = as.factor(id),
                        item = as.factor(item), 
                        answer = as.factor(as.numeric(answer == 'True'))) %>%
  drop_na()

sc_feat_item_clean %>% head()

```

## Exploratory analyses
### Correlations of item level aggregates
```{r}

sc_feat_item_clean %>% group_by(item) %>%
  summarize_if(is.numeric, mean)

sc_feat_item_clean %>% group_by(item) %>%
  summarize_if(is.numeric, mean) %>%
  select_if(is.numeric) %>% cor() %>% round(3) %>%
  as_tibble(rownames = "variables")

```

### Correlations of person level aggregates
```{r}

sc_feat_item_clean %>% group_by(id) %>% 
  summarize_if(is.numeric, mean)

sc_feat_item_clean %>% group_by(id) %>% 
  summarize_if(is.numeric, mean) %>% 
  select_if(is.numeric) %>% 
  cor() %>% round(3) %>% 
  as_tibble(rownames = 'features')

```



# Predict Overall Scores
## Merge data
```{r}

sc_feat_all <- sc_feat %>% inner_join(df_all, by="id") 
sc_feat_all

```


## Exploratory analyses
```{r}

# correlation table 
sc_feat_all_cor <- sc_feat_all %>% dplyr::select(-id) %>% 
  cor() %>% 
  as.data.frame() %>% 
  rownames_to_column() %>% 
  filter(rowname %in% c("pc", "mr", "sc", "VKMR", "SPM")) %>%
  as_tibble(rownames = 'features')

sc_feat_all_cor

```


## Predict SC Scores From Aggregate Process Data
```{r}

sc_feat_all_pred_sc <- sc_feat_all %>% 
  dplyr::select(-c(pc, mr, VKMR, SPM, xPos_range, ray_x_range)) %>% 
  column_to_rownames("id")

cv_results_sc <- nested_cv_reg(data = sc_feat_all_pred_sc, 
                               target = 'sc',
                               outer_method = 'monte_carlo',
                               inner_number = 10, 
                               outer_number = 30,
                               p_mc = 0.85,
                               random_seed = 1)

cv_results_sc$enet %>% as.data.frame() %>% map_dbl(mean)
cv_results_sc$rf %>% as.data.frame() %>% map_dbl(mean)
```


## Predict MR Scores From Aggregate Process Data
```{r}

sc_feat_all_pred_mr <- sc_feat_all %>% 
  dplyr::select(-c(pc, sc, SPM, VKMR, xPos_range, ray_x_range)) %>% 
  column_to_rownames("id")

cv_results_mr <- nested_cv_reg(data = sc_feat_all_pred_mr, 
                           target = 'mr',
                           #outer_method = 'monte_carlo',
                           inner_number = 10, 
                           outer_number = 30,
                           p_mc = 0.85,
                           random_seed = 1)

cv_results_mr$enet %>% as.data.frame() %>% replace( is.na(.), 0) %>% map_dbl(mean)
cv_results_mr$rf %>% as.data.frame() %>% replace( is.na(.), 0) %>% map_dbl(mean)

```


## Predict PC Scores From Aggregate Process Data
```{r}

sc_feat_all_pred_pc <- sc_feat_all %>% 
  dplyr::select(-c(mr, sc, SPM, VKMR, xPos_range, ray_x_range)) %>% 
  column_to_rownames("id")

cv_results_pc <- nested_cv_reg(data = sc_feat_all_pred_pc, 
                               target = 'pc',
                               outer_method = 'monte_carlo',
                               inner_number = 10, 
                               outer_number = 30,
                               p_mc = 0.85,
                               random_seed = 1)

cv_results_pc$enet %>% as.data.frame() %>% replace( is.na(.), 0) %>% map_dbl(mean)
cv_results_pc$rf %>% as.data.frame() %>% replace( is.na(.), 0) %>% map_dbl(mean)

```


## Predict VKMR Scores From Aggregate Process Data
```{r}

sc_feat_all_pred_vkmr <- sc_feat_all %>% 
  dplyr::select(-c(pc, mr, sc, SPM, xPos_range, ray_x_range)) %>% 
  column_to_rownames("id")

cv_results_vkmr <- nested_cv_reg(data = sc_feat_all_pred_vkmr,
                                 target = 'VKMR',
                                 outer_method = 'monte_carlo',
                                 inner_number = 10, 
                                 outer_number = 30,
                                 p_mc = 0.85,
                                 random_seed = 1)

cv_results_vkmr$enet %>% as.data.frame() %>% map_dbl(mean)
cv_results_vkmr$rf %>% as.data.frame() %>% map_dbl(mean)

```


## Predict SPM Scores From Aggregate Process Data
```{r}

sc_feat_all_pred_spm <- sc_feat_all %>%
  dplyr::select(-c(pc, mr, sc, VKMR, xPos_range, ray_x_range)) %>%
  column_to_rownames("id")

cv_results_spm <- nested_cv_reg(data = sc_feat_all_pred_spm,
                                target = 'SPM',
                                outer_method = 'monte_carlo',
                                inner_number = 10,
                                outer_number = 30,
                                p_mc = 0.85,
                                random_seed = 1)

cv_results_spm$enet %>% as.data.frame() %>% replace( is.na(.), 0) %>% map_dbl(mean)
cv_results_spm$rf %>% as.data.frame() %>% replace( is.na(.), 0) %>% map_dbl(mean)

cv_results_spm$enet %>% as.data.frame() %>% map_dbl(mean, na.rm=T)
cv_results_spm$rf %>% as.data.frame() %>% map_dbl(mean, na.rm=T)


```


```{r}

results <- list(cv_results_sc, cv_results_pc, cv_results_mr, cv_results_vkmr, cv_results_spm)
targets <- c('SC', 'PC', 'MR', 'VKMR', 'SPM')

enet_means <- NULL
rf_means <- NULL

for(i in results){
enet_mean <- i$enet %>% as.data.frame() %>% replace( is.na(.), 0) %>% map_dbl(mean)
enet_means <- enet_means %>% rbind(enet_mean)

rf_mean <- i$rf %>% as.data.frame() %>% replace( is.na(.), 0) %>% map_dbl(mean)
rf_means <- rf_means %>% rbind(rf_mean)
}

rownames(enet_means) <-targets
enet_means %>% as.data.frame %>% select(RMSE, MAE, cor_acc)

rownames(rf_means) <-targets
rf_means %>% as.data.frame %>% select(RMSE, MAE, Rsquared, rsq, cor_acc)

```

```{r}

results <- list(cv_results_sc, cv_results_pc, cv_results_mr, cv_results_vkmr, cv_results_spm)
targets <- c('SC', 'PC', 'MR', 'VKMR', 'SPM')

enet_means <- NULL
rf_means <- NULL

for(i in results){
enet_mean <- i$enet %>% as.data.frame() %>% map_dbl(mean, na.rm=T)
enet_means <- enet_means %>% rbind(enet_mean)

rf_mean <- i$rf %>% as.data.frame() %>% map_dbl(mean, na.rm=T)
rf_means <- rf_means %>% rbind(rf_mean)
}

rownames(enet_means) <-targets
enet_means %>% as.data.frame %>% select(RMSE, MAE, Rsquared, cor_acc)

rownames(rf_means) <-targets
rf_means %>% as.data.frame %>% select(RMSE, MAE, Rsquared, rsq, cor_acc) %>% xtable::xtable()

```



# Cluster Analysis of Solution Strategies 
### Identify clusters using UMAP and DBSCAN
```{r}
# select only features
sc_feat_all_feats <- sc_feat_all %>% 
  select(distance_travelled:looking_ratio)

# project features to umap space
umap_config <- umap.defaults
umap_config$random_state <- 123

sc_feat_umap <- umap(sc_feat_all_feats, umap_config, verbose=T)
sc_feat_umap

# run dbscan cluster analysis
dbscan_cluster = dbscan(sc_feat_umap$layout, eps=3)
cluster <- dbscan_cluster$cluster

sc_feat_all_clust <- sc_feat_all %>% 
  cbind(cluster)

plot(sc_feat_umap$layout, col = cluster, xlab = '', ylab = '')
legend("bottom", legend=levels(as.factor(cluster)), pch=16, col=unique(cluster))

ggplot()
```

### Test for cluster differences 
```{r}
# Test for group differences between clusters (ANOVA)
sc_feat_all_clust %>% group_by(cluster) %>% summarize(mean(sc))
anova_results <- aov(sc~as.factor(cluster), data = sc_feat_all_clust)
anova_results %>% summary()
lm(sc~as.factor(cluster), data = sc_feat_all_clust) %>% summary()

sc_feat_all_clust %>% group_by(cluster) %>% summarize(mean(pc))
anova_results <- aov(pc~as.factor(cluster), data = sc_feat_all_clust)
anova_results %>% summary()

sc_feat_all_clust %>% group_by(cluster) %>% summarize(mean(mr))
anova_results <- aov(mr~as.factor(cluster), data = sc_feat_all_clust)
anova_results %>% summary()

sc_feat_all_clust %>% group_by(cluster) %>% summarize(mean(VKMR))
anova_results <- aov(VKMR~as.factor(cluster), data = sc_feat_all_clust)
anova_results %>% summary()

sc_feat_all_clust %>% group_by(cluster) %>% summarize(mean(SPM))
anova_results <- aov(SPM~as.factor(cluster), data = sc_feat_all_clust)
anova_results %>% summary()

# Pairwise t-test
pairwise.t.test(sc_feat_all_clust$sc, sc_feat_all_clust$cluster)
pairwise.t.test(sc_feat_all_clust$pc, sc_feat_all_clust$cluster)
pairwise.t.test(sc_feat_all_clust$mr, sc_feat_all_clust$cluster)
pairwise.t.test(sc_feat_all_clust$VKMR, sc_feat_all_clust$cluster)
pairwise.t.test(sc_feat_all_clust$SPM, sc_feat_all_clust$cluster)

```

### Explore different strategies 
```{r}

sc_feat_all_clust %>% group_by(cluster) %>% summarize(mean(sc))
sc_feat_all_clust %>% filter(cluster==3) %>% map(mean) %>% unlist()

```


```{r}

sc_feat_all_clust %>% filter(cluster!=3) %>% map(mean) %>% unlist()

```

